---
title: "workout01-kimberly-chua.Rmd"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Workout 01: GSW Shot Charts

General Instructions
• Create a folder (i.e. subdirectory) workout01 which will contain all the other subdirectories
and files produced for this HW.
• The folder workout01 (and its contents) will have to be part of a git repository in
your computer, that is associated to the github classroom repository for Stat 133.
Don’t worry about how to do this; we’ll provide more information in lab, and in some
supporting documentation.
• Create a README.md file and include a clear description of what the HW is about.
• Create a folder data which will contain the data files. 
• Create a folder code which will contain the R script files.
• Create a folder output which will contain some R outputs.
• Create a folder images which will contain some plot images.
• Create a folder report which will contain the files for your dynamic document (e.g.
Rmd and derived files).
• In the yaml header of the Rmd file, set the output field as output: github_document
(Do NOT use the default "output: html_document").
• Name your Rmd file as workout01-first-last.Rmd, where first and last are your
first and last names (e.g. workout01-gaston-sanchez.Rmd).
• Use Git to add and commit the changes as you progress with your HW.
• And don’t forget to push your commits to your github repository; you should push the
Rmd and md files, as well as the generated folder and files containing the plot images
and other outputs.
• Submit the link of your repository to bCourses. Do NOT submit any files (we will
actually turn off the uploading files option).
• No html files will be taken into account (no exceptions).

2) Data
The data for this assignment involves the following five CSV files available inside the data/
folder of the github repo https://github.com/ucb-stat133/stat133-hws.
• andre-iguodala.csv
• draymond-green.csv
• kevin-durant.csv
• klay-thompson.csv
• stephen-curry.csv
```{r eval=FALSE}
curl -O https://raw.githubusercontent.com/ucb-stat133/stat133-hws/master/data/andre-iguodala.csv
curl -O https://raw.githubusercontent.com/ucb-stat133/stat133-hws/master/data/draymond-green.csv
curl -O https://raw.githubusercontent.com/ucb-stat133/stat133-hws/master/data/kevin-durant.csv
curl -O https://raw.githubusercontent.com/ucb-stat133/stat133-hws/master/data/klay-thompson.csv
curl -O https://raw.githubusercontent.com/ucb-stat133/stat133-hws/master/data/stephen-curry.csv
curl -O https://raw.githubusercontent.com/ucb-stat133/stat133-hws/master/data/nba2018-players-dictionary.md
```
##2.1) Download the data
• You will need to get a copy of the data files to your local repository.
• The instructions to download the data file should NOT be part of your report
##2.2) Data Dictionary
Create a data dictionary—using markdown syntax—in a separate text file: data-dictionary.md.
Include names of the variables, and a short description. Most of the variable names are
self-descriptive: e.g. team_name, game_date. However, depending on how much you know
about basketball, some variables may be a bit cryptic. So here’s a description for some of
them:
• period: an NBA game is divided in 4 periods of 12 mins each. For example, a value
for period = 1 refers to the first period (the first 12 mins of the game).
• minutes_remaining and seconds_remaining have to do with the amount of time in
minutes and seconds, respectively, that remained to be played in a given period.
• shot_made_flag indicates whether a shot was made (y) or missed (n).
• action_type has to do with the basketball moves used by players, either to pass by
defenders to gain access to the basket, or to get a clean pass to a teammate to score a
two pointer or three pointer.
• shot_type indicates whether a shot is a 2-point field goal, or a 3-point field goal.
• shot_distance: distance to the basket (measured in feet).
• x and y refer to the court coordinates (measured in inches) where a shot occurred .
If you are interested about the dimensions of an NBA basketball court visit these links:
https://www.sportsknowhow.com/basketball/dimensions/nba-basketball-court-dimensions.
html
http://www.sportscourtdimensions.com/wp-content/uploads/2015/02/nba_court_
dimensions_h.png

##3) Data Preparation
The first stage of the assignment has to do with the so-called data preparation phase. The
primary goal of this stage is to create a csv data file shots-data.csv that will contain the
required variables to be used in the visualization phase.
All the R code to complete the data preparation stage must be written in an .R script file
(do NOT confuse with an Rmd file). Name the R script file as make-shots-data-script.R
and save it inside the code/ folder. Include a header (but NOT a yaml header) in the file
containing:
• title: short title
• description: a short description of what the script is about
• input(s): what are the inputs required by the script?
• output(s): what are the outputs created when running the script?
As mentioned above, the raw data for this assignment consists of five data CSV files (one for
each player). To create a main/global table, write code in your R script to carry out these
steps:
• Read in the five data sets, using relative file paths; the decision of the data types for
each column is up to you.
# example: reading file with relative path
curry <- read.csv("../data/stephen-curry.csv", stringsAsFactors = FALSE)
• Add a column name to each imported data frame, that contains the name of the
corresponding player:
– Andre Iguodala
– Graymond Green
– Kevin Durant
– Klay Thompson
– Stephen Curry
• Change the original values of shot_made_flag to more descriptive values: replace "n"
with "shot_no", and "y" with "shot_yes". Hint: you can use logical subsetting for
this operation (no need to use programming structures that we haven’t covered yet).
• Add a column minute that contains the minute number where a shot occurred. For
instance, if a shot took place during period = 1 and minutes_remaining = 8, then
this should correspond to a value minute = 4. Likewise, if a shot took place during
period = 4 and minutes_remaining = 2 then this should correspond to a value
minute = 46. Hint: you can use logical subsetting for these operations (no need to use
programming structures that we haven’t covered yet).
• Use sink() to send the summary() output of each imported data frame into individuals
text files: andre-iguodala-summary.txt, draymond-green-summary.txt, etc. During
each sinking operation, the produced summaries should be sent to the output/
folder using relative paths.
• Use the row binding function rbind() to stack the tables into one single data frame
(or tibble object).
• Export (i.e. write) the assembled table as a CSV file shots-data.csv inside the folder
data/. Use a relative path for this operation.
• Use sink() to send the summary() output of the assembled table. Send this output to
a text file named shots-data-summary.txt inside the output/ folder. Use a relative
path when exporting the R output.
4) Shot Charts
This part of the assignment has to do with the creation of shot charts. Write the code in
an R script called make-shot-charts-script.R, and save it in the code/ folder. Make sure
you include a header with fields about title, description, inputs, and outputs.
Let’s begin by describing some preliminary steps to make this type of charts. In this example,
I will use the data for Klay Thompson. Suppose we have a data.frame (or tibble) klay with
Klay Thompson’s data. We can get a scatterplot of the shots using the x and y variables
(i.e. coordinates of the shots):
```{r}
# scatterplot
klay_scatterplot <- ggplot(data = klay) +
geom_point(aes(x = x, y = y, color = shot_made_flag))
```
To make the shot chart more meaningful, and visually appealing, we will add a background
with the image of an NBA basketball court. The JPG file with this image is inside the folder
images/ of the github repo https://github.com/ucb-stat133/stat133-hws/tree/master/image.
You need to import the file nba-court.jpg to the images/ folder of your HW directory.
Likewise, you will need the following R packages: "jpeg" and "grid".
```{r}
# court image (to be used as background of plot)
court_file <- "../images/nba-court.jpg"
# create raste object
court_image <- rasterGrob(
readJPEG(court_file),
width = unit(1, "npc"),
height = unit(1, "npc"))
```
What’s happening? The function readJPEG() allows you to import the .jpg file. In turn,
rasterGrob() takes the .jpg file and converts it into a raster graphical object. This object
will be used as the background for the ggplot graphic:
```{r}
# shot chart with court background
klay_shot_chart <- ggplot(data = klay) +
annotation_custom(court_image, -250, 250, -50, 420) +
geom_point(aes(x = x, y = y, color = shot_made_flag)) +
ylim(-50, 420) +
ggtitle('Shot Chart: Klay Thompson (2016 season)') +
theme_minimal()
```
##4.1) Shot charts of each player
Write code in the R script make-shot-charts-script.R to create shot charts (with court
backgrounds) for each player, and save the plots in PDF format, with dimensions width =
6.5 and height = 5 inches, inside the folder images/:
• andre-iguodala-shot-chart.pdf
• draymond-green-shot-chart.pdf
• kevin-durant-shot-chart.pdf
• klay-thompson-shot-chart.pdf
• stephen-curry-shot-chart.pdf
##4.2) Facetted Shot Chart
Create one graph, using facetting, to show all the shot charts in one image. A hypothetical
facetted graphic is shown below.
image dimensions width = 8 and height = 7 inches.
Likewise, save this image in PNG format as gsw-shot-charts.png, inside the folder images/.
Specify image dimensions width = 8 and height = 7 inches.
Feel free to choose different visual attributes for the geometric object points.
##5) Report
Imagine that you are hired by an Oakland-based sports media company to write an article
about the shooting statistics of the five GSW players: Iguodala, Green, Durant, Thompson,
and Curry.
They want the article to contain at least:
• three tables about Effective Shooting Percentage (see section 5.1 below).
• the PNG image of the facetted shot charts (the one obtained in section 4.2)
Rmd file settings
• Your Rmd file must be located in the report/ subdirectory.
• Recall that you need to specify the output field in the yaml header as github_document
(NOT an html_document). See image below.
• To insert an image (in a given file), you can use include_graphics(). For example,
suppose you have a file named 'image1.png' in the working directory of your Rmd file;
here’s how you could import that image in your report:
##5.1) Effective Shooting Percentage
Use the data in shots-data.csv to create three tables to summarize Effective Shooting
percentages by player (import the data in your Rmd file using a relative path).
All three tables should have the format of the following depicted diagram:
2PT Effective Shooting % by Player: 2 PT Field Goal effective shooting percentage by
player, arranged in descending order by percentage.
3PT Effective Shooting % by Player: 3 PT Field Goal effective shooting percentage by
player, arranged in descending order by percentage.
Effective Shooting % by Player: Overall (i.e. including 2PT and 3PT Field Goals)
effective shooting percentage by player, arranged in descending order by percentage.
##5.2) Narrative
Here’s a list of general specifications for the text in your article:
• Introduction: Write a clear and captivating introduction that grabs the reader’s
attention. And describe the purpose of the report.
• Organize the content: Organize the content of your report in different sections:
e.g. motivation, background, data, analysis, discussion, conclusions, and references.
• Images: You may include some images like pictures about players, or other figures in
addition to the requested facetted shot chart.
• Code: You may also include code snippets (e.g. code chunks) with output that supports
your narrative.
• Take-home message: Make sure to include a summarizing statement of the content
in your article. If the reader had to remember one major thing from your article, what
would that be?
• Length: How long should your article be? There is not really a unique answer for this
question. It depends on what you decide to write about. Try writing a post no shorter
than 1,000 words, and preferably no longer than 5,000 words (these values are just a
recommendation).
To give you some inspiration, here are some real articles about GSW players (from different
seasons) looking at different performance aspects:
• https://on.nba.com/2Vv1t2u
• https://bit.ly/2HcDJMD
• https://bit.ly/2BWTmoc
• https://bit.ly/2EnDkUD